Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Mechanical Perturbations

### 2024-06-24 main.tex
Small updates in text

\subsection{Perturbations}
%
We apply external longitudinal forces to the ends of each handlebar using an
adapted Bump'Em system~\cite{Tan2020} which is designed with four motors
working in tandem, Figure~\ref{fig:setup-diagram}. The four motors are
programmed to apply a light force at all times to keep the ropes taught and to
track a commanded force profile using a PID controller running on a
microprocessor (Arduino Mega 2560, Italy). We measure the force applied by each
motor at the handlebar via four inline load cells, each with a max load of
250~\si{\newton}. One load cell was damaged during pilot experiments, and is
replaced by a load cell with a max load of 500~\si{\newton}. The commanded
force profiles are designed to apply an external pulsive torque to the front
assembly (handlebars, forks, wheel) at magnitudes varying from 20 to 200. The
four motors are arranged at the four corners of a 1~\si{\meter} wide treadmill
(EC-90 Flat, Maxon Group, Switzerland) that can reach speeds of 18~\si{\kilo\meter\per\hour}.
The general design is described in detail in the Van De Velde's MSc
hesis~\citep{vandeVelde2022a} and the physical arrangement  is shown in
Figure~\ref{fig:participant-in-set-up}. Our modifications relative to that
described in the thesis include simplifying the controller with a inexpensive
microcontroller and the use of a simpler safety harness.
Figure~\ref{fig:perturbation_0} shows an example perturbation.
%
% NOTE : I get a compile error with the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[width=75mm]{figures/participant-in-set-up.png}}
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[width=65mm]{figures/setup-diagram.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations.}
  \label{fig:experimental-setup}
\end{figure}

@@

Subjects wore a helmet and fall safety harness attached to the ceiling
Figure~\ref{fig:participant-in-set-up}. We allowed them to practice riding on
the treadmill until they indicated they were comfortable enough to have
perturbations applied. For most, this was less than a 10~\si{\minute} warm up.
We then asked the rider to ride for 90~\si{\second}, attempting to maintain the
location of their front wheel on the center line of the treadmill. \todo{The results 
from this are not included because the balance-assist was turned on for some participants}
We define a
``fall'' on the treadmill by two criteria: the rider removes their foot from
the pedal and places it on the ground or the bicycle wheel exceeds the width of
the treadmill. We then applied perturbations in random directions (clockwise or
counter clockwise), starting at 20~\si{\newton} and increasing the magnitude by 
20~\si{\newton\meter} until the participants falls.
Figure~\ref{fig:perturbation-sequence} shows an example result of a
perturbation.  We log the magnitude that causes the first fall to characterize
that subject's nominal fall threshold. Following this we apply 20
perturbations of random magnitudes to the cyclist while they ride at a constant
speed and record which perturbations cause a fall. We let the cyclist rest and
then perform another 20 perturbations. We randomize whether the balance assist
system is on or off during the first or second set of perturbations.
%
\begin{figure}
  \centering
  \subcaptionbox{Start}{\includegraphics[width=40mm]{figures/start.png}}
  \subcaptionbox{During}{\includegraphics[width=40mm]{figures/during.png}}
  \subcaptionbox{After}{\includegraphics[width=40mm]{figures/after.png}}
  \subcaptionbox{Recovery}{\includegraphics[width=40mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}

>>>>

\subsection{Perturbations}
%
We apply external longitudinal forces to the ends of each handlebar using an
adapted Bump'Em system~\cite{Tan2020} which is designed with four motors
working in tandem, Figure~\ref{fig:setup-diagram}. The four motors are
programmed to apply a light force at all times to keep the ropes taught and to
track a commanded force profile using a PID controller running on a
microprocessor (Arduino Mega 2560, Italy). We measure the force applied by each
motor at the handlebar via four inline load cells, each with a max load of
250~\si{\newton}. The left rear load cell  was damaged during pilot experiments, and is
replaced by a load cell with a max load of 500~\si{\newton}. The commanded
force profiles are designed to apply an external pulsive torque to the front
assembly (handlebars, forks, wheel) at magnitudes varying from 20 to 200~\si{\Newton} per motor. 
This results in a torque between 16 and 160~\si{\Newton\meter}. The
four motors are arranged at the four corners of a 1~\si{\meter} wide treadmill
(EC-90 Flat, Maxon Group, Switzerland) that can reach speeds of 18~\si{\kilo\meter\per\hour}.
The general design is described in detail in the Van De Velde's MSc
hesis~\citep{vandeVelde2022a} and the physical arrangement  is shown in
Figure~\ref{fig:participant-in-set-up}. Our modifications relative to that
described in the thesis include simplifying the controller with a inexpensive
microcontroller and the use of a simpler safety harness.
Figure~\ref{fig:perturbation_0} shows an example perturbation.
%
% NOTE : I get a compile error with the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[width=75mm]{figures/participant-in-set-up.png}}
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[width=65mm]{figures/setup-diagram.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations.}
  \label{fig:experimental-setup}
\end{figure}

@@

Subjects wore a helmet and fall safety harness attached to the ceiling
Figure~\ref{fig:participant-in-set-up}. We allowed them to practice riding on
the treadmill until they indicated they were comfortable enough to have
perturbations applied. For most, this was less than a 10~\si{\minute} warm up.
We then asked the rider to ride for 90~\si{\second}, attempting to maintain the
location of their front wheel on the center line of the treadmill. \todo{The results 
from this are not included because the balance-assist was turned on for some participants}
We define a
``fall'' on the treadmill by two criteria: the rider removes their foot from
the pedal and places it on the ground or the bicycle wheel exceeds the width of
the treadmill. We then applied perturbations in random directions (clockwise or
counter clockwise), starting at 20~\si{\newton} and increasing the magnitude by 
30~\si{\newton\meter} until the participants falls.
Figure~\ref{fig:perturbation-sequence} shows an example result of a
perturbation.  We log the magnitude that causes the first fall to characterize
that subject's nominal fall threshold. Following this we apply 20
perturbations of varying magnitudes to the cyclist while they ride at a constant
speed and record which perturbations cause a fall. We let the cyclist rest and
then perform another 20 perturbations. We randomize whether the balance assist
system is on or off during the first or second set of perturbations.

The magnitude of the perturbation is determined using an adaptive staircase procedure. 
\todo{Cite to correct paper from thesis}
The adaptive staircase takes the nominal fall threshold as input and randomly chooses 
one of five possible forces that are centred around the nominal fall threshold with 
10~\si{\Newton} intervals. If the perturbation results in a fall, the nominal fall 
threshold is decreased with 10~\si{\Newton}. If the perturbation does not results in a 
fall, the nominal fall threshold is increased with 10~\si{\Newton}. This adaptive
staircase procedure ensures that approximately 50\% of the perturbations applied to the
rider result in a fall. 
%
\begin{figure}
  \centering
  \subcaptionbox{Start}{\includegraphics[width=40mm]{figures/start.png}}
  \subcaptionbox{During}{\includegraphics[width=40mm]{figures/during.png}}
  \subcaptionbox{After}{\includegraphics[width=40mm]{figures/after.png}}
  \subcaptionbox{Recovery}{\includegraphics[width=40mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}
